Parallel multi-objective optimization using co-evolutionary islands with overlapping search space

### 2016-12-18 hpmoon-coor.tex
I don't like "coevolutionary" or "adaptive"
It's simply partial evolution and variable. Why do you call it coevolutionary? Refs #12

\title{Parallel multi-objective optimization using co-evolutionary
  islands with overlapping search space}
% New title suggestion

@@

The application of MOEAs to solve real-world problems have gained attention in the last years \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress}. These kind of problems requires a high number of variables, meaning that the EAs need to deal with large individuals and spent significant extra time for crossover, mutation and migration. Different authors have proposed methods to divide the decision space (the chromosome) to improve the performance and solution quality. In this aspect, the co-evolution model is a dimension-distributed model where a high-dimensional problem is divided into lower dimensional problems \citep{Gong15models,Tonda12cooperative}, and evolved separately. One example of application of this technique was described in \citep{Kimovski15Parallel}. The method presented in that works involves different workers that evolve sub-populations created and recombined by a master process, which performs different recombination alternatives from the parts returned by worker processes. A high dimensional problem was used to compare these
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
coevolutionary island model was used.  Although in both previous works
significant speedups were attained, only a low number of nodes were
used (8). However, when increasing the number of islands, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

In our preliminary work \citep{Garcia16hpmoon} we tried to demonstrate
% tried to or actually did it? What did we prove? 
that,  instead of making every island focus in a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome in a coevolutionary multi-objective island-based algorithm
can improve the quality of the solutions in the same computing time
when the number of islands increases. We discovered that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. Moreover, the experiments
were conducted in a mono-processor and synchronous island model. This
motivates us to continue this research line; in this paper we propose
a new method that automatically adapts the size of the overlapped
sections. In order to test the validity of the method we have made
tests in a real cluster
with up to 128 nodes, with extra number of islands. % We have to
                                % enphatize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ

@@

\section{Coevolutionary algorithms compared}
\label{sec:coevo}

@@

As in the baseline (B), an individual is migrated to another random island after a fixed number of generations. In the new island, this individual will be considered as one of the others in the island, crossed and mutated in the same way, depending on the island identifier. Note that, on the contrary of other works such as the ones described by Talbi et al. \citep{Talbi08Parallel} all the islands deal with complete chromosomes for fitness calculation, so our approach can deal with no decomposable problems. The proposed methods for our experimental section are as follows:

@@

\subsection{Adaptive coevolution with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but
instead using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands. As a first
approach to automatically calculate this value the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were $>$8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Figure \ref{fig:adaptive} explains this method.

>>>>

\title{Parallel multi-objective optimization using co-evolutionary
  islands with overlapping search space}
% New title suggestion
% Should coevolutionary go somewhere? Are they coevolutionary? - jj

@@

The application of MOEAs to solve real-world problems have gained attention in the last years \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress}. These kind of problems requires a high number of variables, meaning that the EAs need to deal with large individuals and spent significant extra time for crossover, mutation and migration. Different authors have proposed methods to divide the decision space (the chromosome) to improve the performance and solution quality. In this aspect, the co-evolution model is a dimension-distributed model where a high-dimensional problem is divided into lower dimensional problems \citep{Gong15models,Tonda12cooperative}, and evolved separately. One example of application of this technique was described in \citep{Kimovski15Parallel}. The method presented in that works involves different workers that evolve sub-populations created and recombined by a master process, which performs different recombination alternatives from the parts returned by worker processes. A high dimensional problem was used to compare these
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
coevolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

In our preliminary work \citep{Garcia16hpmoon} we tried to demonstrate
% tried to or actually did it? What did we prove? 
that,  instead of making every island focus in a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome in a coevolutionary multi-objective island-based algorithm
can improve the quality of the solutions in the same computing time
when the number of islands increases. We discovered that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. Moreover, the experiments
were conducted in a mono-processor and synchronous island model. This
motivates us to continue this research line; in this paper we propose
a new method that automatically adapts the size of the overlapped
sections. In order to test the validity of the method we have made
tests in a real cluster
with up to 128 nodes, with extra number of islands. % We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ

@@

\section{Coevolutionary algorithms compared} % This should be called
                                % methodology. The object of the
                                % section is to explain the
                                % methodology used to test the
                                % problem, in order to do that we have
                                % tested different versions that prove
                                % that our methodology works for the
                                % precise reasons we think it works- JJ
\label{sec:coevo}

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier. Note that, differently from other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel} all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ

The methods we have tested are presented in the next subsections.

@@

\subsection{Adaptive coevolution with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but
instead using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ
As a first
approach to automatically calculate this value the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were $>$8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Figure \ref{fig:adaptive} explains this method.
